---
title: "Logistic Regression"
author: "Emily Van Buren"
date: "`r Sys.Date()`"
output: html_document
---

# Set Up Working Directory 

In this markdown we will go over how to perform parallel processed logistic regression models for all genes expressed in the SCTLD/WP classification study using [RegParallel](https://github.com/kevinblighe/RegParallel). 

```{r, eval=FALSE}
### load data 
library(readr)
library(tidyr)
library(dplyr)
library(mixOmics)
library(ggrepel)
library(PCAtools)
library("ggalt")
library(DESeq2)
library(sva)
library(glmnet)
library(rIsing)
library(RegParallel)
library(caret)
library(pROC)

# set working directory 
setwd("~/Documents/Documents/UTA/RESEARCH/coral_classification/disease_classification/LG/")

# set up variables - not batched

# SCTLD Run 
gene_counts <- as.data.frame(read.csv("~/Documents/Documents/UTA/RESEARCH/coral_classification/disease_classification/Normalization_DESeq/normalized_counts_vstCounts_Diseased.csv", row.names = "Entry"))

# set working directory 

metadata <- as.data.frame(read.csv("~/Documents/Documents/UTA/RESEARCH/coral_classification/disease_classification/counts/metadata_diseased.csv", row.names = "Sample"))
metadata_diseased <- metadata %>% filter(metadata$Training_Testing == c("Training"))
metadata_diseased$Training_Testing

colnames(gene_counts) == rownames(metadata_diseased)
```

# SCTLD Training 

In this round, we will be looking for genes predictive individually for SCTLD lesions only. There will be three types of algorithms tested, a generalized linear model with binomial, generalized linear model with glm.fit, and a linear model. 

## Set Variables 
```{r, eval=FALSE}
# Change to binary values - SCTLD first 
metadata_diseased <- metadata_diseased %>%
  mutate(Disease = ifelse(Disease == "SCTLD",1,0))

rlddata_coral <- as.data.frame(t(gene_counts))
rlddata_coral$SCTLD <- metadata_diseased$Disease

dim(rlddata_coral)
# [1]    34 18598
```

## Generalized Linear Model with binomial(link = 'logit') 

The first RegParallel was run as a generalized linear model with binomial(link = 'logit') link. The 18598 vst-normalized genes were input individually against the formula SCTLD ~ [gene_ID]. Outputs of this include beta, standard error, Z-score, P-value, Odds Ratio and the Odds Ratio Upper and lower values. Models were filtered for p-value < 0.05. In this model type, 1865 genes models were associated with SCTLD. 

```{r, eval=FALSE}
### Res1 - Generalized Linear Model with binomial(link = 'logit')

#check a subset 

res1 <- RegParallel(
  data = rlddata_coral[ ,1:18598],
  formula = 'SCTLD ~ [*]',
  FUN = function(formula, data)
    glm(formula = formula,
        data = data,
        family = binomial(link = 'logit')),
  FUNtype = 'glm',
  variables = colnames(rlddata_coral)[1:18597])

# ##############################
# #RegParallel
# ##############################
# 
# System is:
#   -- Darwin
# Blocksize:
#   -- 500
# Cores / Threads:
#   -- 3
# Terms included in model:
#   -- SCTLD
# First 5 formulae:
#   -- SCTLD ~ A0A023I7E1
# -- SCTLD ~ A0A044RE18
# -- SCTLD ~ A0A061IR73
# -- SCTLD ~ A0A067XMP1
# -- SCTLD ~ A0A075F932
# Done!

res1_df <- as.data.frame(res1[order(res1$P, decreasing=FALSE),])
subset_res1 <- as.data.frame(subset(res1_df, P<0.05))
# 1865 genes Associated with SCTLD
write.csv(subset_res1, file = "RegParallel_res1_SCTLD.csv")
```

### Area Under the Reciever Operating Characteristics (ROC) Curve (AUC) 

AUC curves were calculated for each model under the training dataset for those with a p-value < 0.05. A loop was generated to test the gene models, and then AUC was filtered >75. A total of 250 genes had an AUC of > 75. 
```{r, eval=FALSE}
#### AUC-ROC check 

rlddata_subset1 <- rlddata_coral[,c(subset_res1$Variable)]
rlddata_subset1$SCTLD <- rlddata_coral$SCTLD 

# create a loop
pdf(file = "ROC_SCTLD_res1_batched.pdf", height = 8, width = 8)
mod_summaries_res1_all <- list() 
for (i in colnames(rlddata_subset1)){
  roc_x <- roc(rlddata_subset1$SCTLD, 
               rlddata_subset1[,i],percent=TRUE, 
               partial.auc=c(100, 90), partial.auc.correct=TRUE,
               partial.auc.focus="sens", ci=TRUE, boot.n=100, 
               ci.alpha=0.9, stratified=FALSE, plot=TRUE, col= 'red')
  mod_summaries_res1_all[i] <-print(roc_x$auc)
  
}
dev.off()
lapply(mod_summaries_res1_all, function(x) write.table( data.frame(x), 'mod_summaries_res1_SCTLD.csv', append= T, sep=',' ))
# in excel use find and select (go to special) to remove empty spaces 
# use variables from summary to define the Variables 
aucs <-  as.data.frame(read.csv("mod_summaries_res1_SCTLD.csv"))
aucs_noNAs <- na.omit(aucs)
nrow(aucs_noNAs[aucs_noNAs$AUC > 50,])
# [1] 1771
nrow(aucs_noNAs[aucs_noNAs$AUC > 75,])
# [1] 250
nrow(aucs_noNAs[aucs_noNAs$AUC > 80,])
# [1] 46
nrow(aucs_noNAs[aucs_noNAs$AUC > 85,])
# [1] 15
```

### Transcriptome Presence Absense Filtering 
Finally, genes were then filtered for presence in four or more transcriptomes using the previously generated Presence Absence file. 

```{r, eval=FALSE}
# filer out top annotating genes & annotate 
PA <- as.data.frame(read.csv("/Volumes/My Passport/coral_classification/transcriptomes/annotations/PA_7sp.csv"))
aucs_noNAs_PA <- merge(aucs_noNAs,PA,by="Entry")
aucs_noNAs_PA <- aucs_noNAs_PA[ !duplicated(aucs_noNAs_PA$Entry), ]
nrow(aucs_noNAs_PA[aucs_noNAs_PA$Total > 3,])
# [1] 1216
aucs_noNAs_PA <- aucs_noNAs_PA %>% filter(Total > 3)
uniprot <- read.csv("/Volumes/My Passport/coral_classification/transcriptomes/annotations/uniprot_7species_reviewed_yes.csv")
aucs_noNAs_PA <- merge(aucs_noNAs_PA,uniprot, by="Entry")
write.csv(aucs_noNAs_PA, file = "res1_SCTLD_best_predictive.csv")
```

## Generlized Linear model with glm.fit

The second RegParallel was run as a generalized linear model with with glm.fit. The 18598 vst-normalized genes were input individually against the formula SCTLD ~ [gene_ID]. Outputs of this include beta, standard error, Z-score, P-value, Odds Ratio and the Odds Ratio Upper and lower values. Models were filtered for p-value < 0.05. In this model type, 4094 genes models were associated with SCTLD. 

```{r, eval=FALSE}
res2 <- RegParallel(
  data = rlddata_coral,
  formula = '[*] ~ SCTLD',
  FUN = function(formula, data)
    glm(formula = formula,
        data = data,
        method = 'glm.fit'),
  FUNtype = 'glm',
  variables = colnames(rlddata_coral)[1:18597],
  p.adjust = "none")
# ##############################
# #RegParallel
# ##############################
# 
# System is:
#   -- Darwin
# Blocksize:
#   -- 500
# Cores / Threads:
#   -- 3
# Terms included in model:
#   -- SCTLD
# First 5 formulae:
#   -- A0A023I7E1 ~ SCTLD
# -- A0A044RE18 ~ SCTLD
# -- A0A061IR73 ~ SCTLD
# -- A0A067XMP1 ~ SCTLD
# -- A0A075F932 ~ SCTLD
# Done!

subset_res2 <- as.data.frame(subset(res2, P<0.05))
# 4094 Genes Associated with WP
write.csv(subset_res2, file = "RegParallel_res2_SCTLD.csv")
```

### Area Under the Reciever Operating Characteristics (ROC) Curve (AUC) 

AUC curves were calculated for each model under the training dataset for those with a p-value < 0.05. A loop was generated to test the gene models, and then AUC was filtered >75. A total of 250 genes had an AUC of > 75. 
```{r, eval=FALSE}
rlddata_subset2 <- rlddata_coral[,c(subset_res2$Variable)]
rlddata_subset2$SCTLD <- rlddata_coral$SCTLD 

# create a loop
pdf(file = "ROC_SCTLD_res2.pdf", height = 8, width = 8)
mod_summaries_res2_all <- list() 
for (i in colnames(rlddata_subset2)){
  roc_x <- roc(rlddata_subset2$SCTLD, 
               rlddata_subset2[,i],percent=TRUE, 
               partial.auc=c(100, 90), partial.auc.correct=TRUE,
               partial.auc.focus="sens", ci=TRUE, boot.n=100, 
               ci.alpha=0.9, stratified=FALSE, plot=TRUE, col= 'red')
  mod_summaries_res2_all[i] <-print(roc_x$auc)
  
}
dev.off()
lapply(mod_summaries_res2_all, function(x) write.table( data.frame(x), 'mod_summaries_res2_SCTLD.csv', append= T, sep=',' ))
# in excel use find and select (go to special) to remove empty spaces 
# use variables from summary to define the Variables 
aucs <-  as.data.frame(read.csv("mod_summaries_res2_SCTLD.csv"))
aucs_noNAs <- na.omit(aucs)
nrow(aucs_noNAs[aucs_noNAs$AUC > 50,])
# [1] 1771
nrow(aucs_noNAs[aucs_noNAs$AUC > 75,])
# [1] 250
nrow(aucs_noNAs[aucs_noNAs$AUC > 80,])
# [1] 46
nrow(aucs_noNAs[aucs_noNAs$AUC > 85,])
# [1] 15
```

### Transcriptome Presence Absense Filtering 
Finally, genes were then filtered for presence in four or more transcriptomes using the previously generated Presence Absence file. 

```{r, eval=FALSE}
# filer out top annotating genes & annotate 
PA <- as.data.frame(read.csv("/Volumes/My Passport/coral_classification/transcriptomes/annotations/PA_7sp.csv"))
aucs_noNAs_PA <- merge(aucs_noNAs,PA,by="Entry")
aucs_noNAs_PA <- aucs_noNAs_PA[ !duplicated(aucs_noNAs_PA$Entry), ]
nrow(aucs_noNAs_PA[aucs_noNAs_PA$Total > 3,])
# [1] 1216
aucs_noNAs_PA <- aucs_noNAs_PA %>% filter(Total > 3)
uniprot <- read.csv("/Volumes/My Passport/coral_classification/transcriptomes/annotations/uniprot_7species_reviewed_yes.csv")
aucs_noNAs_PA <- merge(aucs_noNAs_PA,uniprot, by="Entry")
write.csv(aucs_noNAs_PA, file = "res2_SCTLD_best_predictive.csv")
```

## Linear Model  

The third RegParallel was run as a linear model. The 18598 vst-normalized genes were input individually against the formula SCTLD ~ [gene_ID]. Outputs of this include beta, standard error, Z-score, P-value, Odds Ratio and the Odds Ratio Upper and lower values. Models were filtered for p-value < 0.05. In this model type, 4094 genes models were associated with SCTLD. 

```{r, eval=FALSE}
res3 <- RegParallel(
  data = rlddata_coral,
  formula = '[*] ~ SCTLD',
  FUN = function(formula, data)
    lm(formula = formula,
       data = data),
  FUNtype = 'lm',
  variables = colnames(rlddata_coral)[1:18597],
  p.adjust = "none")

# ##############################
# #RegParallel
# ##############################
# 
# System is:
#   -- Darwin
# Blocksize:
#   -- 500
# Cores / Threads:
#   -- 3
# Terms included in model:
#   -- SCTLD
# First 5 formulae:
#   -- A0A023I7E1 ~ SCTLD
# -- A0A044RE18 ~ SCTLD
# -- A0A061IR73 ~ SCTLD
# -- A0A067XMP1 ~ SCTLD
# -- A0A075F932 ~ SCTLD
# Done!

subset_res3 <- as.data.frame(subset(res3, P<0.05))
# 4094 Genes Associated with WP
write.csv(subset_res3, file = "RegParallel_res3_SCTLD.csv")
```

### Area Under the Reciever Operating Characteristics (ROC) Curve (AUC) 

AUC curves were calculated for each model under the training dataset for those with a p-value < 0.05. A loop was generated to test the gene models, and then AUC was filtered >75. A total of 398 genes had an AUC of > 75. 
```{r, eval=FALSE}
rlddata_subset3 <- rlddata_coral[,c(subset_res3$Variable)]
rlddata_subset3$SCTLD <- rlddata_coral$SCTLD 

# create a loop
pdf(file = "ROC_SCTLD_res3.pdf", height = 8, width = 8)
mod_summaries_res3_all <- list() 
for (i in colnames(rlddata_subset3)){
  roc_x <- roc(rlddata_subset3$SCTLD, 
               rlddata_subset3[,i],percent=TRUE, 
               partial.auc=c(100, 90), partial.auc.correct=TRUE,
               partial.auc.focus="sens", ci=TRUE, boot.n=100, 
               ci.alpha=0.9, stratified=FALSE, plot=TRUE, col= 'red')
  mod_summaries_res3_all[i] <-print(roc_x$auc)
  
}
dev.off()
lapply(mod_summaries_res3_all, 
       function(x) write.table( data.frame(x), 
                                'mod_summaries_res3_SCTLD.csv', append= T, sep=',' ))
aucs <-  as.data.frame(read.csv("mod_summaries_res3_SCTLD.csv"))
aucs_noNAs <- na.omit(aucs)
nrow(aucs_noNAs[aucs_noNAs$AUC > 50,])
# [1] 2450
nrow(aucs_noNAs[aucs_noNAs$AUC > 75,])
# [1] 398
nrow(aucs_noNAs[aucs_noNAs$AUC > 80,])
# [1] 54
nrow(aucs_noNAs[aucs_noNAs$AUC > 85,])
# [1] 18
```

### Transcriptome Presence Absense Filtering 
Finally, genes were then filtered for presence in four or more transcriptomes using the previously generated Presence Absence file. 

```{r, eval=FALSE}
# filer out top annotating genes & annotate 
PA <- as.data.frame(read.csv("/Volumes/My Passport/coral_classification/transcriptomes/annotations/PA_7sp.csv"))
aucs_noNAs_PA <- merge(aucs_noNAs,PA,by="Entry")
aucs_noNAs_PA <- aucs_noNAs_PA[ !duplicated(aucs_noNAs_PA$Entry), ]
nrow(aucs_noNAs_PA[aucs_noNAs_PA$Total > 3,])
# [1] 1543
aucs_noNAs_PA <- aucs_noNAs_PA %>% filter(Total > 3)
uniprot <- read.csv("/Volumes/My Passport/coral_classification/transcriptomes/annotations/uniprot_7species_reviewed_yes.csv")
aucs_noNAs_PA <- merge(aucs_noNAs_PA,uniprot, by="Entry")
write.csv(aucs_noNAs_PA, file = "res3_SCTLD_best_predictive.csv")
```

## Save Data 
```{r, eval=FALSE}
save(gene_counts, i, metadata, metadata_diseased, mod_summaries_res2_all, mod_summaries_res3_all, res2, res3, rlddata_coral, rlddata_subset2, rlddata_subset3, roc_x, subset_res1, subset_res2, subset_res3, file = "SCTLD_LG.RData")
```

# WP Training 

In this round, we will be looking for genes predictive individually for WP lesions only. There will be three types of algorithms tested, a generalized linear model with binomial,  

## Set Variables 
```{r, eval=FALSE}
# set working directory 
metadata <- as.data.frame(read.csv("/Volumes/My Passport/coral_classification/disease_classification/counts/metadata_diseased.csv", row.names = "Sample"))
metadata_diseased <- metadata %>% filter(metadata$Training_Testing == c("Training"))
metadata_diseased$Training_Testing

colnames(gene_counts) == rownames(metadata_diseased)

# Change to binary values - WP first 
metadata_diseased <- metadata_diseased %>%
  mutate(Disease = ifelse(Disease == "WP",1,0))

rlddata_coral <- as.data.frame(t(gene_counts))
rlddata_coral$WP <- metadata_diseased$Disease

dim(rlddata_coral)
#[1]    34 18598
```

## Generalized Linear Model with binomial(link = 'logit') 

The first RegParallel was run as a generalized linear model with binomial(link = 'logit') link. The 18598 vst-normalized genes were input individually against the formula WP ~ [gene_ID]. Outputs of this include beta, standard error, Z-score, P-value, Odds Ratio and the Odds Ratio Upper and lower values. Models were filtered for p-value < 0.05. In this model type, 1865 genes models were associated with WP. 

```{r, eval=FALSE}
res1 <- RegParallel(
  data = rlddata_coral[ ,1:18598],
  formula = 'WP ~ [*]',
  FUN = function(formula, data)
    glm(formula = formula,
        data = data,
        family = binomial(link = 'logit')),
  FUNtype = 'glm',
  variables = colnames(rlddata_coral)[1:18597])
# ##############################
# #RegParallel
# ##############################
# 
# System is:
#   -- Darwin
# Blocksize:
#   -- 500
# Cores / Threads:
#   -- 3
# Terms included in model:
#   -- WP
# First 5 formulae:
#   -- WP ~ A0A023I7E1
# -- WP ~ A0A044RE18
# -- WP ~ A0A061IR73
# -- WP ~ A0A067XMP1
# -- WP ~ A0A075F932
# Done!

res1_df <- as.data.frame(res1[order(res1$P, decreasing=FALSE),])
subset_res1 <- as.data.frame(subset(res1_df, P<0.05)) 
# 1865 genes signficiantly associated with WP
write.csv(subset_res1, file = "RegParallel_res1_WP.csv")
```

### Area Under the Reciever Operating Characteristics (ROC) Curve (AUC) 

AUC curves were calculated for each model under the training dataset for those with a p-value < 0.05. A loop was generated to test the gene models, and then AUC was filtered >75. A total of 54 genes had an AUC of > 75. 
```{r, eval=FALSE}
rlddata_subset1 <- rlddata_coral[,c(subset_res1$Variable)]
rlddata_subset1$WP <- rlddata_coral$WP

# create a loop
pdf(file = "ROC_WP_res1.pdf", height = 8, width = 8)
mod_summaries_res1_all <- list() 
for (i in colnames(rlddata_subset1)){
  roc_x <- roc(rlddata_subset1$WP, 
               rlddata_subset1[,i],percent=TRUE, 
               partial.auc=c(100, 90), partial.auc.correct=TRUE,
               partial.auc.focus="sens", ci=TRUE, boot.n=100, 
               ci.alpha=0.9, stratified=FALSE, plot=TRUE, col= 'red')
  mod_summaries_res1_all[i] <-print(roc_x$auc)
  
}
dev.off()
lapply(mod_summaries_res1_all, function(x) write.table( data.frame(x), 'mod_summaries_res1_WP.csv', append= T, sep=',' ))
# in excel use find and select (go to special) to remove empty spaces 
# use variables from summary to define the Variables 
aucs <-  as.data.frame(read.csv("mod_summaries_res1_WP.csv"))
aucs_noNAs <- na.omit(aucs)
nrow(aucs_noNAs[aucs_noNAs$AUC > 50,])
# [1] 1788
nrow(aucs_noNAs[aucs_noNAs$AUC > 75,])
# [1] 54
nrow(aucs_noNAs[aucs_noNAs$AUC > 80,])
# [1] 13
nrow(aucs_noNAs[aucs_noNAs$AUC > 85,])
# [1] 5
```

### Transcriptome Presence Absense Filtering 
Finally, genes were then filtered for presence in four or more transcriptomes using the previously generated Presence Absence file. 

```{r, eval=FALSE}
# filer out top annotating genes & annotate 
PA <- as.data.frame(read.csv("/Volumes/My Passport/coral_classification/transcriptomes/annotations/PA_7sp.csv"))
aucs_noNAs_PA <- merge(aucs_noNAs,PA,by="Entry")
aucs_noNAs_PA <- aucs_noNAs_PA[ !duplicated(aucs_noNAs_PA$Entry), ]
nrow(aucs_noNAs_PA[aucs_noNAs_PA$Total > 3,])
# [1] 1227
aucs_noNAs_PA <- aucs_noNAs_PA %>% filter(Total > 3)
uniprot <- read.csv("/Volumes/My Passport/coral_classification/transcriptomes/annotations/uniprot_7species_reviewed_yes.csv")
aucs_noNAs_PA <- merge(aucs_noNAs_PA,uniprot, by="Entry")
nrow(aucs_noNAs_PA[aucs_noNAs_PA$AUC > 75,])
# 52 genes present in 4 or more species with AUC 75 or more
write.csv(aucs_noNAs_PA, file = "res1_WP_best_predictive.csv")
```

## Generlized Linear model with glm.fit

The second RegParallel was run as a generalized linear model with with glm.fit. The 18598 vst-normalized genes were input individually against the formula A0A023I7E1 ~ WP. Outputs of this include beta, standard error, Z-score, P-value, Odds Ratio and the Odds Ratio Upper and lower values. Models were filtered for p-value < 0.05. In this model type, 4094 genes models were associated with WP. 

```{r, eval=FALSE}
res2 <- RegParallel(
  data = rlddata_coral,
  formula = '[*] ~ WP',
  FUN = function(formula, data)
    glm(formula = formula,
        data = data,
        method = 'glm.fit'),
  FUNtype = 'glm',
  variables = colnames(rlddata_coral)[1:18597],
  p.adjust = "none")

# ##############################
# #RegParallel
# ##############################
# 
# System is:
#   -- Darwin
# Blocksize:
#   -- 500
# Cores / Threads:
#   -- 3
# Terms included in model:
#   -- WP
# First 5 formulae:
#   -- A0A023I7E1 ~ WP
# -- A0A044RE18 ~ WP
# -- A0A061IR73 ~ WP
# -- A0A067XMP1 ~ WP
# -- A0A075F932 ~ WP
# Done!

subset_res2 <- as.data.frame(subset(res2, P<0.05))
# 4094 Genes Associated with WP
write.csv(subset_res2, file = "RegParallel_res2_WP.csv")
```

### Area Under the Reciever Operating Characteristics (ROC) Curve (AUC) 

AUC curves were calculated for each model under the training dataset for those with a p-value < 0.05. A loop was generated to test the gene models, and then AUC was filtered >75. A total of 56 genes had an AUC of > 75. 
```{r, eval=FALSE}
rlddata_subset2 <- rlddata_coral[,c(subset_res2$Variable)]
rlddata_subset2$WP <- rlddata_coral$WP

# create a loop
pdf(file = "ROC_WP_res2.pdf", height = 8, width = 8)
mod_summaries_res2_all <- list() 
for (i in colnames(rlddata_subset2)){
  roc_x <- roc(rlddata_subset2$WP, 
               rlddata_subset2[,i],percent=TRUE, 
               partial.auc=c(100, 90), partial.auc.correct=TRUE,
               partial.auc.focus="sens", ci=TRUE, boot.n=100, 
               ci.alpha=0.9, stratified=FALSE, plot=TRUE, col= 'red')
  mod_summaries_res2_all[i] <-print(roc_x$auc)
  
}
dev.off()
lapply(mod_summaries_res2_all, function(x) write.table( data.frame(x), 'mod_summaries_res2_WP.csv', append= T, sep=',' ))
# in excel use find and select (go to special) to remove empty spaces 
# use variables from summary to define the Variables 
aucs <-  as.data.frame(read.csv("mod_summaries_res2_WP.csv"))
aucs_noNAs <- na.omit(aucs)
nrow(aucs_noNAs[aucs_noNAs$AUC > 50,])
# [1] 3869
nrow(aucs_noNAs[aucs_noNAs$AUC > 75,])
# [1] 56
nrow(aucs_noNAs[aucs_noNAs$AUC > 80,])
# [1] 13
nrow(aucs_noNAs[aucs_noNAs$AUC > 85,])
# [1] 5
```

### Transcriptome Presence Absense Filtering 
Finally, genes were then filtered for presence in four or more transcriptomes using the previously generated Presence Absence file. 

```{r, eval=FALSE}
PA <- as.data.frame(read.csv("/Volumes/My Passport/coral_classification/transcriptomes/annotations/PA_7sp.csv"))
aucs_noNAs_PA <- merge(aucs_noNAs,PA,by="Entry")
aucs_noNAs_PA <- aucs_noNAs_PA[ !duplicated(aucs_noNAs_PA$Entry), ]
nrow(aucs_noNAs_PA[aucs_noNAs_PA$Total > 3,])
# [1] 3
aucs_noNAs_PA <- aucs_noNAs_PA %>% filter(Total > 3)
uniprot <- read.csv("/Volumes/My Passport/coral_classification/transcriptomes/annotations/uniprot_7species_reviewed_yes.csv")
aucs_noNAs_PA <- merge(aucs_noNAs_PA,uniprot, by="Entry")
nrow(aucs_noNAs_PA[aucs_noNAs_PA$AUC > 75,])
# 53 genes present in 4 or more species with AUC 75 or more
write.csv(aucs_noNAs_PA, file = "res2_WP_best_predictive.csv")
```

## Linear Model  

The third RegParallel was run as a linear model. The 18598 vst-normalized genes were input individually against the formula [gene_ID] ~ WP. Outputs of this include beta, standard error, Z-score, P-value, Odds Ratio and the Odds Ratio Upper and lower values. Models were filtered for p-value < 0.05. In this model type, 4094 genes models were associated with WP. 

```{r, eval=FALSE}
res3 <- RegParallel(
  data = rlddata_coral,
  formula = '[*] ~ WP',
  FUN = function(formula, data)
    lm(formula = formula,
       data = data),
  FUNtype = 'lm',
  variables = colnames(rlddata_coral)[1:18597],
  p.adjust = "none")
# ##############################
# #RegParallel
# ##############################
# 
# System is:
#   -- Darwin
# Blocksize:
#   -- 500
# Cores / Threads:
#   -- 3
# Terms included in model:
#   -- WP
# First 5 formulae:
#   -- A0A023I7E1 ~ WP
# -- A0A044RE18 ~ WP
# -- A0A061IR73 ~ WP
# -- A0A067XMP1 ~ WP
# -- A0A075F932 ~ WP
# Done!

subset_res3 <- as.data.frame(subset(res3, P<0.05))
# 4094 Genes Associated with WP
write.csv(subset_res3, file = "RegParallel_res3_WP.csv")
```

### Area Under the Reciever Operating Characteristics (ROC) Curve (AUC) 

AUC curves were calculated for each model under the training dataset for those with a p-value < 0.05. A loop was generated to test the gene models, and then AUC was filtered >75. A total of 398 genes had an AUC of > 75. 


```{r, eval=FALSE}
rlddata_subset3 <- rlddata_coral[,c(subset_res3$Variable)]
rlddata_subset3$WP <- rlddata_coral$WP 

# create a loop
pdf(file = "ROC_WP_res3.pdf", height = 8, width = 8)
mod_summaries_res3_all <- list() 
for (i in colnames(rlddata_subset3)){
  roc_x <- roc(rlddata_subset3$WP, 
               rlddata_subset3[,i],percent=TRUE, 
               partial.auc=c(100, 90), partial.auc.correct=TRUE,
               partial.auc.focus="sens", ci=TRUE, boot.n=100, 
               ci.alpha=0.9, stratified=FALSE, plot=TRUE, col= 'red')
  mod_summaries_res3_all[i] <-print(roc_x$auc)
  
}
dev.off()
lapply(mod_summaries_res3_all, 
       function(x) write.table( data.frame(x), 
                                'mod_summaries_res3_WP.csv', append= T, sep=',' ))
aucs <-  as.data.frame(read.csv("mod_summaries_res3_WP.csv"))
aucs_noNAs <- na.omit(aucs)
nrow(aucs_noNAs[aucs_noNAs$AUC > 50,])
# [1] 3869
nrow(aucs_noNAs[aucs_noNAs$AUC > 75,])
# [1] 56
nrow(aucs_noNAs[aucs_noNAs$AUC > 80,])
# [1] 13
nrow(aucs_noNAs[aucs_noNAs$AUC > 85,])
# [1] 5
```

### Transcriptome Presence Absense Filtering 
Finally, genes were then filtered for presence in four or more transcriptomes using the previously generated Presence Absence file. 
```{r, eval=FALSE}
PA <- as.data.frame(read.csv("/Volumes/My Passport/coral_classification/transcriptomes/annotations/PA_7sp.csv"))
aucs_noNAs_PA <- merge(aucs_noNAs,PA,by="Entry")
aucs_noNAs_PA <- aucs_noNAs_PA[ !duplicated(aucs_noNAs_PA$Entry), ]
nrow(aucs_noNAs_PA[aucs_noNAs_PA$Total > 3,])
# [1] 1579
aucs_noNAs_PA <- aucs_noNAs_PA %>% filter(Total > 3)
uniprot <- read.csv("/Volumes/My Passport/coral_classification/transcriptomes/annotations/uniprot_7species_reviewed_yes.csv")
aucs_noNAs_PA <- merge(aucs_noNAs_PA,uniprot, by="Entry")
nrow(aucs_noNAs_PA[aucs_noNAs_PA$AUC > 75,])
# 53 genes present in 4 or more species with AUC 75 or more
write.csv(aucs_noNAs_PA, file = "res3_WP_best_predictive.csv")
```

## Save Data 
```{r, eval=FALSE}
save(gene_counts, i, metadata, metadata_diseased, mod_summaries_res1_all, mod_summaries_res2_all, mod_summaries_res3_all, res1, res2, res3, rlddata_coral, rlddata_subset1, rlddata_subset2, rlddata_subset3, roc_x, subset_res1, subset_res2, subset_res3, file = "WP_LG.RData")
```

# Venn Diagrams 

To compare the genes found in each of the model training, we will use venn diagrams to visualize and obtain overlapping genes if applicable. 

## set working directory 

```{r, eval=FALSE}
library(tidyverse)
library(ggvenn)
library(ggplot2)
library(dplyr)
library(ggpubr)

setwd("/Volumes/My Passport/coral_classification/disease_classification/LG/")

WP_res1 <- as.data.frame(read.csv("res1_WP_best_predictive.csv"))
WP_res1 <- WP_res1 %>% filter(AUC > 75)
WP_res2 <- as.data.frame(read.csv("res2_WP_best_predictive.csv"))
WP_res2 <- WP_res2 %>% filter(AUC > 75)
WP_res3 <- as.data.frame(read.csv("res3_WP_best_predictive.csv"))
WP_res3 <- WP_res3 %>% filter(AUC > 75)

SCTLD_res1 <- as.data.frame(read.csv("res1_SCTLD_best_predictive.csv"))
SCTLD_res1 <- SCTLD_res1 %>% filter(AUC > 75)
SCTLD_res2 <- as.data.frame(read.csv("res2_SCTLD_best_predictive.csv"))
SCTLD_res2 <- SCTLD_res2 %>% filter(AUC > 75)
SCTLD_res3 <- as.data.frame(read.csv("res3_SCTLD_best_predictive.csv"))
SCTLD_res3 <- SCTLD_res3 %>% filter(AUC > 75)
```

## Unique Function Set Up 

Functions made here include interset, union, and setdiff, which can help identify unique sections of comparisons of lists. 

```{r, eval=FALSE}
## Isolation of Intersect/Unique Genes in Venn Diagram

Intersect <- function (x) {  
  # Multiple set version of intersect
  # x is a list
  if (length(x) == 1) {
    unlist(x)
  } else if (length(x) == 2) {
    intersect(x[[1]], x[[2]])
  } else if (length(x) > 2){
    intersect(x[[1]], Intersect(x[-1]))
  }
}

Union <- function (x) {  
  # Multiple set version of union
  # x is a list
  if (length(x) == 1) {
    unlist(x)
  } else if (length(x) == 2) {
    union(x[[1]], x[[2]])
  } else if (length(x) > 2) {
    union(x[[1]], Union(x[-1]))
  }
}

Setdiff <- function (x, y) {
  # Remove the union of the y's from the common x's. 
  # x and y are lists of characters.
  xx <- Intersect(x)
  yy <- Union(y)
  setdiff(xx, yy)
}
```

## WP Comparisons 

This venn diagram compares the genes significant to the three trials of RegParallel. 
```{r, eval=FALSE}
# WP venn diagrams 
WP_LG <- list(
  res1 = WP_res1$Entry,
  res2 = WP_res2$Entry,
  res3 = WP_res3$Entry
)

# Venn Diagram 
ggvenn(
  WP_LG,
  fill_color = c("#FFF7BC", "#A1DAB4", "#41B6C4"),
  stroke_size = 0.5, set_name_size = 5
)
```

## SCTLD Comparisons 
This venn diagram compares the genes significant to the three trials of RegParallel. 
```{r, eval=FALSE}
# SCTLD venn diagrams 
SCTLD_LG <- list(
  res1 = SCTLD_res1$Entry,
  res2 = SCTLD_res2$Entry,
  res3 = SCTLD_res3$Entry
)

# Venn Diagram 
ggvenn(
  SCTLD_LG,
  fill_color = c("#FFF7BC", "#A1DAB4", "#41B6C4"),
  stroke_size = 0.5, set_name_size = 5
)
```

## WP and SCTLD Comparisons 

We will be continuing with binomial logistic regression as the choice algorithm for this study. We created a venn diagram to compare the overlap of signficiant genes from the WP and SCTLD studies. We also generate a list of the genes into excel sheets. 

### Binomial 
```{r, eval=FALSE}
# Binomial Logistic Regression venn diagrams 
res1_LG <- list(
  WP = WP_res1$Entry,
  SCTLD = SCTLD_res1$Entry
)

# Venn Diagram 
ggvenn(
  res1_LG,
  fill_color = c("#d8b365", "#5ab4ac"),
  stroke_size = 0.5, set_name_size = 5
)
```

### Shared Biomarkers 

```{r, eval=FALSE}
# shared biomarkers 
overlap_biomarkers <- Intersect(res1_LG)

# Unique biomarkers
uniprot <- read.csv("~/Documents/Documents/UTA/RESEARCH/coral_classification/transcriptomes/annotations/uniprot_7species_reviewed_yes.csv", row.names = "Entry")
SCTLD_unique <- Setdiff(res1_LG[c("SCTLD")], res1_LG[c("WP")])
SCTLD_unique_anno <- uniprot[SCTLD_unique,]
SCTLD_unique_anno$Entry <- rownames(SCTLD_unique_anno)
SCTLD_unique_anno <- merge(SCTLD_unique_anno,SCTLD_res1,by="Entry")
write.csv(SCTLD_unique_anno, file = "~/Documents/Documents/UTA/RESEARCH/coral_classification/disease_classification/LG/SCTLD_unique_biomarkers_annot.csv")
WP_unique <- Setdiff(res1_LG[c("WP")], res1_LG[c("SCTLD")])
WP_unique_anno <- uniprot[WP_unique,]
WP_unique_anno$Entry <- rownames(WP_unique_anno)
WP_unique_anno <- merge(WP_unique_anno,WP_res1,by="Entry")
write.csv(WP_unique_anno, file = "~/Documents/Documents/UTA/RESEARCH/coral_classification/disease_classification/LG/WP_unique_biomarkers_annot.csv")
```

## Save Data

```{r, eval=FALSE}
ls()
# [1] "Intersect"          "overlap_biomarkers" "res1_LG"            "SCTLD_LG"          
# [5] "SCTLD_res1"         "SCTLD_res2"         "SCTLD_res3"         "SCTLD_unique"      
# [9] "Setdiff"            "Union"              "WP_LG"              "WP_res1"           
# [13] "WP_res2"            "WP_res3"            "WP_unique"   
save(Intersect, overlap_biomarkers, res1_LG, SCTLD_LG, SCTLD_res1, SCTLD_res2, 
     SCTLD_res3, SCTLD_unique, Setdiff, Union, WP_LG, WP_res1, WP_res2, 
     WP_res3, WP_unique, file = "venn_diagrams.RData")
```
